Acceleration of Derivative Calculations with Application to Radial Basis Function -- Finite-Differences on the Intel MIC Architecture

### 2014-04-16 acm/acm_paper_submitted.tex
acm_paper_final_truetype.pdf and acm_paper_submitted.tex are the sources submitted ot the ACM conference (camera ready). (and dummy.ps)

\begin{figure}[t]
\centering
\includegraphics[width=0.60\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
\caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square. In nonsparse format, the DM  is $1024\times1024$ with 32 nonzero entries per row.}
\end{figure}

@@

\ger{The main idea of this paper, computing 16 derivatives at a time,
can be applied on different architectures such as multiprocessor
multicore or GPU. To illustrate this fact, we conducted similar experiments
on Cascade at the Minnesota Supercomputing Institute
using two Intel Xeon E5 2670 nodes (8 cores, 2 threads and 20 Mbytes of L2
cache per core, with hyperthreading disabled.) The cores support the AVX instruction set. 
Both the compiler-generated code and the code written with the 
help of the AVX instruction set reach a read bandwidth of
85~GB/s. Using cache lines of 512 bits, we find a best/worst speeds of 2.4/20\; Gflops/s and 
32/120\; Gflops/s, respectively for the 1vec/1mat and 4vec/4mat cases. 
Our implementation on Cascade reaches 45 Gflops for the 4/4\, case, or 45/120=37\% of peak theoretical 
performance on this particular processor setup. For references, we achieve 
140/210=66\% of peak theoretical performance on the MIC. Note that peak performance 
is not that of the processor, but of the best possible algorithm on the processor. 
The gain in the 4/4 case with respect to 4x 1/1 case is respectively 4x on the MIC 
and 2.5x on the front end.  
}

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=0.60\linewidth]{figures/RBFStencil_n32_a.pdf}\label{fig:stencil}
\caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square. In nonsparse format, the DM  is $1024\times1024$ with 32 nonzero entries per row.}
\end{figure}

@@

\ger{The main idea of this paper, computing 16 derivatives at a time,
can be applied on different architectures such as multiprocessor
multicore or GPU. To illustrate this fact, we conducted similar experiments
on Cascade at the Minnesota Supercomputing Institute
using two Intel Xeon E5 2670 nodes (8 cores, 2 threads and 20 Mbytes of L2
cache per core, with hyperthreading disabled.) The cores support the AVX instruction set. 
Both the compiler-generated code and the code written with the 
help of the AVX instruction set reach a read bandwidth of
85~GB/s. Using cache lines of 512 bits, we find a best/worst speeds of 2.4/20\; Gflops/s and 
32/120\; Gflops/s, respectively for the 1vec/1mat and 4vec/4mat cases. 
Our implementation on Cascade reaches 45 Gflops for the 4/4\, case, or 45/120=37\% of peak theoretical 
performance on this particular processor setup. For reference, we achieve 
140/210=66\% of peak theoretical performance on the MIC. Note that peak performance 
is not that of the processor, but of the best possible algorithm on the processor. 
The gain in the 4/4 case with respect to 4x 1/1 case is 4x on the MIC 
and 2.5x on the front end, respectively.  
}
